---
title: "Export data"
subtitle: "ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Lien Reysehove
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Install packages

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(etn)            # To interact with ETN
library(glue)
```

# Connect to database

Get username/password from environment (therefore requires this script to be run on <http://rstudio.lifewatch.be>) and connect to database:

```{r}
my_con <- connect_to_etn(Sys.getenv("username"), Sys.getenv("password"))
```

# Define selection criteria

We will extract the animal and transmitter data based on the animal ID's of interest. These are the animal ID's integrated in the following animal projects:

- 2011 Rivierprik
- 2012 Leopoldkanaal
- 2013 Albertkanaal (excluding _Sync tag_)
- 2014 Demer
- 2015 Dijle
- 2015 Phd Verhelst (only for _Anguilla anguilla_)
- Homarus
- Phd Reubens (only for _Gadus morhua_)

```{r}
get_animals(connection = my_con, 
            animal_project = "2011_rivierprik")
```

