---
title: "Select transmitter and receiver data"
subtitle: "ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Lien Reysehove
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---

# Setup 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Install packages

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(etn)            # To interact with ETN
library(lubridate)      # To process date and time information
library(glue)
```

# Connect to database

Get username/password from environment (therefore requires this script to be run on <http://rstudio.lifewatch.be>) and connect to database:

```{r}
my_con <- connect_to_etn(Sys.getenv("username"), Sys.getenv("password"))
```


# Select Animals data

Extract all information related to the following animal projects:

- 2012 Leopoldkanaal
- 2013 Albertkanaal (excluding Sync tag data)
- 2014 Demer
- 2015 Dijle
- 2015 Phd Verhelst (only for _Anguilla anguilla_)
- Homarus
- Phd Reubens (only for _Gadus morhua_)

```{r}
query <- glue_sql(
    "SELECT *
    FROM vliz.animals_view
    WHERE 
      (projectcode = '2011_rivierprik' 
        OR projectcode = '2012_leopoldkanaal'
        OR projectcode = '2013_albertkanaal'
        OR projectcode = '2014_demer'
        OR projectcode = '2015_dijle'
        OR (projectcode = '2015_phd_verhelst' AND scientific_name = 'Anguilla anguilla')
        OR projectcode = 'homarus'
        OR projectcode = 'phd_reubens')
      AND NOT scientific_name = 'Sentinel'
      AND NOT scientific_name = 'Sync tag'
    " ,
    .con = my_con)

animals <- dbGetQuery(my_con, query)
as_tibble(animals)
```

Save data in `raw`:

```{r}
animals %>% write_csv(here::here("data", "raw", "animals.csv"), na = "")
```

# Select transmitter data

Select `tag_full_id` and save as separate vector:

```{r}
tag_full_id <- animals %>% pull(tag_full_id)
```

Select tag information fomr `vliz_tags`:

```{r}
query <- glue_sql(
    "SELECT tags.*, animals.projectcode, animals.scientific_name
     FROM vliz.tags
        LEFT JOIN vliz.animal_tag_release ON (animal_tag_release.tag_fk = tags.id_pk)
        LEFT JOIN vliz.animals_view animals ON (animals.id_pk = animal_tag_release.animal_fk)
    WHERE 
      (projectcode = '2011_rivierprik' 
        OR projectcode = '2012_leopoldkanaal'
        OR projectcode = '2013_albertkanaal'
        OR projectcode = '2014_demer'
        OR projectcode = '2015_dijle'
        OR (projectcode = '2015_phd_verhelst' AND scientific_name = 'Anguilla anguilla')
        OR projectcode = 'homarus'
        OR projectcode = 'phd_reubens')
      AND NOT scientific_name = 'Sentinel'
      AND NOT scientific_name = 'Sync tag'
    ", 
    .con = my_con)

transmitters <- dbGetQuery(my_con, query)
as_tibble(transmitters)
```

Export transmitters data:

```{r}
transmitters %>% write_csv(here::here("data", "raw", "transmitters.csv"), na = "")
```

# Select deployments

Extract all receivers involved in the detections of the specific animal projects listed earlier. We can't select on specific animal projects here. Hence, we use the network projects included in the animal projects.

```{r}
query <- glue_sql(
    "SELECT *
      FROM vliz.deployments_view AS deployments
      WHERE projectcode = 'albert'
        OR projectcode = 'bovenschelde'
        OR projectcode = 'bpns'
        OR projectcode = 'demer'
        OR projectcode = 'dijle'
        OR projectcode = 'leopoldkanaal'
        OR projectcode = 'maas'
        OR projectcode = 'saeftinghe'
        OR projectcode = 'ws1'
        OR projectcode = 'ws2'
        OR projectcode = 'ws3'    ",
    .con = my_con)

deployments <- dbGetQuery(my_con, query)
as_tibble(deployments)
```

Export `deployments` as .csv:

```{r}
write_csv(deployments, "../data/raw/deployments.csv", na = "")
```

# Select receivers

```{r}
query <- glue_sql("
      SELECT DISTINCT 
        receivers.* ,
        deployments.projectcode,
        etn_group.name as owner_organisation
      FROM vliz.receivers
        JOIN vliz.deployments_view deployments ON (receivers.id_pk = deployments.receiver_fk)
        JOIN vliz.etn_group AS etn_group ON (receivers.owner_group_fk = etn_group.id_pk)
       WHERE projectcode = 'albert'
        OR projectcode = 'bovenschelde'
        OR projectcode = 'bpns'
        OR projectcode = 'demer'
        OR projectcode = 'dijle'
        OR projectcode = 'leopoldkanaal'
        OR projectcode = 'maas'
        OR projectcode = 'saeftinghe'
        OR projectcode = 'ws1'
        OR projectcode = 'ws2'
        OR projectcode = 'ws3'    ",
      .con = my_con)

receivers <- dbGetQuery(my_con, query)
as_tibble(receivers)
```

Export `receivers` as .csv:

```{r}
write_csv(receivers, "../data/raw/receivers.csv", na = "")
```

