---
title: "Darwin Core mapping"
subtitle: "ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Lien Reysehove
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
---

This document describes how the data from the [European Tracking Network (etn)](http://www.lifewatch.be/etn/) are standardized to the Darwin Core format.
The mapping is based on the ongoing work in the [Darwin Core for biologging](https://github.com/tdwg/dwc-for-biologging/), following the specifications for the [Ocean Tracking Network (OTN)](https://github.com/tdwg/dwc-for-biologging/wiki/Acoustic-sensor-enabled-tracking-of-blue-sharks). The data are mapped to an [Event Core](http://rs.gbif.org/core/dwc_event_2016_06_21.xml), which is linked to an [Occurrence extension](http://rs.gbif.org/core/dwc_event_2016_06_21.xml) and [Extended Measurements or Facts extension](http://rs.gbif.org/extension/obis/extended_measurement_or_fact.xml) .

- **Event core**: includes the capture, surgery and release event of the specific animals, as well as the deployments of the receiver network
- **Occurrence extension**: includes the capture occurrence of specific animals, as well as the detections of each of these animals by the receivers in the network.
- **Extended measurement or facts extension**: measurements or facts related to the animal surgery event (transmitter specifications) or receiver deployment event, as wel as measurements or facts related to the capture occurrence and detection occurrence.

The basic elements of this mapping scripts are a series of [SQL scripts](https://github.com/inbo/etn-occurrences/tree/master/sql) generated to extract and transform the data from the database.

# Setup 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Install packages

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(DBI)            # To connect to database
library(etn)            # To interact with ETN
library(glue)           # To write SQL with variables
```

# Connect to database

Get username/password from environment (therefore requires this script to be run on <http://rstudio.lifewatch.be>) and connect to database:

```{r}
my_con <- connect_to_etn(Sys.getenv("username"), Sys.getenv("password"))
```

# Select relevant data

1. Select animals:

```{r}
animal_ids <- read_csv(here::here("data", "interim", "animal_ids.csv"))
animal_ids <- pull(animal_ids)
```

2. Select network projects:

```{r}
network_projects <- read_csv(here::here("data", "interim", "network_projects.csv"))
network_projects <- pull(network_projects)
```

# Query data

## Event data

### Animal event

Query animal event data (including the capture, surgery and release event):

```{r}
event_animals_sql <- glue_sql(read_file(here::here("sql", "dwc_event_animals.sql")), .con = my_con)
event_animals_data <- dbGetQuery(my_con, event_animals_sql)
```

Export data as .csv:

```{r}
write_csv(event_animals_data, here::here("data","processed", "dwc_event_animals.csv"), na = "")
```

### Receiver deployment event

Query receiver deployment event data:

```{r}
event_receiver_deployments_sql <- glue_sql(read_file(here::here("sql", "dwc_event_receiver_deployments.sql")), .con = my_con)
event_receiver_deployments_data <- dbGetQuery(my_con, event_receiver_deployments_sql)
```

Export data as.csv:

```{r}
write_csv(event_receiver_deployments_data, here::here("data","processed", "dwc_event_receiver_deployments.csv"), na = "")
```

## Occurrence data

### Animal capture occurrences

Query animal capture occurrence data:

```{r}
occ_animals_sql <- glue_sql(read_file(here::here("sql", "dwc_occ_animals.sql")), .con = my_con)
occ_animals_data <- dbGetQuery(my_con, occ_animals_sql)
```

Export data as .csv:

```{r}
write_csv(occ_animals_data, here::here("data","processed", "dwc_occ_animals.csv"), na = "")
```

### Detections occurrences

The detection occurrence dataset is too large to extract as a whole. Therefore, we use a custom function generated to extract the data for each animal separately.  

1. Load custom function to iterate over animal ids:

```{r}
source(here::here("src", "functions", "download_occ_detections.R"))
```

2. Query and save data (one csv file per animal id):

```{r}
download_occ_detections(
  sql_file = here::here("sql", "dwc_occ_detections.sql"),
  download_directory = here::here("data", "processed", "dwc_occ_detections"),
  animal_ids = animal_ids,
  con = my_con,
  overwrite = TRUE
)
```

## Extended measurements or facts (emof) data

### Emof related to animal capture occurrence:

Query animal emof capture data:

```{r}
emof_animals_capture_sql <- glue_sql(read_file(here::here("sql", "dwc_emof_animals_capture.sql")), .con = my_con)
emof_animals_capture_data <- dbGetQuery(my_con, emof_animals_capture_sql)
```

### Emof related to animal surgery event:

Query animal emof surgery data:

```{r}
emof_animals_surgery_sql <- glue_sql(read_file(here::here("sql", "dwc_emof_animals_surgery.sql")), .con = my_con)
emof_animals_surgery_data <- dbGetQuery(my_con, emof_animals_surgery_sql)
```

### Combine animal emof data

```{r}
emof_animals_data <- 
  emof_animals_capture_data %>% bind_rows(emof_animals_surgery_data)
```

Export data as .csv:

```{r}
write_csv(emof_animals_data, here::here("data","processed", "dwc_emof_animals.csv"), na = "")
```

### Emof related to receiver deployment event

Query data:

```{r get_ref_data}
emof_receiver_deployments_sql <- glue_sql(read_file(here::here("sql", "dwc_emof_receiver_deployments.sql")), .con = my_con)
emof_receiver_deployments_data <- dbGetQuery(my_con, emof_receiver_deployments_sql)
```

Export as .csv:

```{r}
write_csv(emof_receiver_deployments_data, here::here("data","processed", "dwc_emof_receiver_deployments.csv"), na = "")
```

### Emof related to detection occurrences

Query data:

```{r get_ref_data}
emof_detections_sql <- glue_sql(
  read_file(here::here("sql", "dwc_emof_detections.sql")),
  animal_ids = animal_ids,
  .con = my_con
)
emof_detections_data <- dbGetQuery(my_con, emof_detections_sql)
```

Export as .csv:

```{r}
write_csv(emof_detections_data, here::here("data","processed", "dwc_emof_detections.csv"), na = "")
```
