---
title: "Darwin Core mapping"
subtitle: "For: ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

This document describes how we map the checklist data to Darwin Core. The source file for this document can be found [here](https://github.com/inbo/etn-occurrences/blob/master/src/dwc_mapping.Rmd).

# Setup

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}
library(tidyverse) # To transform data
library(magrittr)  # For %<>% pipes
library(janitor)   # To clean input data
```

Set file paths (all paths should be relative to this script):
 
```{r}
# Raw files:
raw_data_file = "../data/raw/etn-dump.csv"

# Processed files:
dwc_machine_occurrence_file = "../data/processed/machine_occurrence.csv"
dwc_human_occurrence_file = "../data/processed/human_occurrence.csv"
dwc_mof_file = "../data/processed/measurementorfact.csv"
```

# Read and pre-process raw data

Create a data frame `raw_data` from the source data:

```{r}
raw_data <- read.csv(raw_data_file, fileEncoding = "UTF-8")
```

Clean the data somewhat:

```{r}
raw_data %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

Show information about data frame:

```{r}
str(raw_data)
```
