---
title: "Darwin Core mapping"
subtitle: "For: ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Peter Desmet
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
---

This document describes how we map the checklist data to Darwin Core. The source file for this document can be found [here](https://github.com/inbo/etn-occurrences/blob/master/src/dwc_mapping.Rmd).

# Setup

```{r setup, echo = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}
library(dplyr)     # To transform data
library(magrittr)  # For %<>% pipes
library(janitor)   # To clean input data
library(lubridate) # To work with dates
```

Set file paths (all paths should be relative to this script):
 
```{r}
# Raw files:
raw_detections_file = "../data/raw/etn_dump.csv"

# Processed files:
dwc_occ_detections_file = "../data/processed/machine_obs_occurrence.csv"
dwc_occ_animal_tagging_file = "../data/processed/human_obs_occurrence.csv"
dwc_mof_file = "../data/processed/measurementorfact.csv"
```

# Read and pre-process raw data

Create a data frame `raw_data` from the source data:

```{r read_raw_data}
raw_detections <- read.csv(raw_detections_file, fileEncoding = "UTF-8")
```

Clean the data somewhat:

```{r}
raw_detections %<>%
  remove_empty("rows") %>%    # Remove empty rows
  clean_names()               # Have sensible (lowercase) column names
```

Show information about data frame:

```{r}
str(raw_detections)
```

Preview data:

```{r}
head(raw_detections)
```

Save columns names

```{r}
raw_detections_colnames <- colnames(raw_detections)
```

## Info per project/species

Show information grouped by project and species:

```{r}
raw_detections %>%
  group_by(animal_project_name, animals_scientific_name) %>%
  summarize(
    detections = n(),
    individuals = n_distinct(animal_id_pk),
    stations = n_distinct(deployment_station_name),
    start_date = round_date(min(as.Date(detections_datetime)), unit = "day"),
    end_date = round_date(max(as.Date(detections_datetime)), unit = "day")
  )
```

# Create occurrence core (detections)

```{r start_detections}
occ_detections <- raw_detections
```

Sort by transmitter and date:

```{r}
occ_detections %<>% arrange(detections_transmitter, detections_datetime)
```

## Term mapping

Map the source data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml) (but in the classic Darwin Core order):

### type

```{r}
occ_detections %<>% mutate(type = "Event")
```

### language

```{r}
occ_detections %<>% mutate(language = "en")
```

### license

```{r}
occ_detections %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

### rightsHolder

The owner of the tag:

```{r}
occ_detections %<>% mutate(rightsHolder = recode(as.character(tag_owner_organization),
  "INBO"                         = "INBO",
  .default = "",
  .missing = ""
))
```

Show mapped values:

```{r}
occ_detections %>% 
  select(tag_owner_organization, rightsHolder) %>%
  group_by_all() %>%
  summarize(records = n()) %>%
  arrange(tag_owner_organization)
```

### accessRights

```{r}
occ_detections %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use")
```

### datasetID

```{r}
occ_detections %<>% mutate(datasetID = "") # TODO: add DOI
```

### institutionCode

Same as the `rightsHolder`:

```{r}
occ_detections %<>% mutate(institutionCode = rightsHolder)
```

### datasetName

```{r}
occ_detections %<>% mutate(datasetName = "Acoustic telemetry data of fish in the Scheldt river basin and the Belgian Part of the North Sea (BPNS)") # TODO: verify
```

### basisOfRecord

```{r}
occ_detections %<>% mutate(basisOfRecord = "MachineObservation")
```

### informationWithheld

```{r eval = FALSE}
occ_detections %<>% mutate(informationWithheld = "see metadata") # TODO: currently not included, to be defined
```

### dynamicProperties

```{r}
occ_detections %<>% mutate(dynamicProperties = paste0(
  "{\"transmitter\":\"", detections_transmitter, "\", ",
  "\"receiver\":\"", detections_receiver, "\"}"
))
```

### occurrenceID

```{r}
occ_detections %<>% mutate(occurrenceID = paste("etn", "lifewatch", detections_id_pk, sep = ":"))
```

Check for duplicate `occurrenceID`s (should be 0):

```{r}
anyDuplicated(occ_detections$occurrenceID)
```

### organismID

```{r eval = FALSE}
occ_detections %<>% mutate(organismID = animals_animal_id) # TODO: This one is often NA. detections_transmitter would have been nicer
```

### eventDate

`datetime` assumed to be UTC. For 3D analyses milliseconds will be required, but these won't be available in the source data until VRL imports are supported:

```{r}
occ_detections %<>% mutate(eventDate = format(as.POSIXct(detections_datetime), format = "%Y-%m-%dT%H:%M:%SZ")) # TODO: verify if UTC
```

### samplingProtocol

```{r}
# TODO: refer to DOI of methodology paper? Can some information be derived from source data?
# receivers_receiver_type: acoustic_telemetry, SVN, active
# receivers_model_number: VR2W, NA, 122325
# animals_capture_method: NA, LINE FISHING, FYKE NETS
```

### locationID

The `deployments_station_name` is a fixed code for that deployment location. It should always be populated. The `detections_receiver` code is not adequate, as a receiver can be moved from one location to another.

```{r}
occ_detections %<>% mutate(locationID = deployment_station_name)
```

Check `locationID` is not `NA` (should be 0):

```{r}
occ_detections %>%
  filter(is.na(locationID)) %>%
  nrow()
```

### decimalLatitude

There are several columns with coordinates information. Of those the **deployment** coordinates of the receiver are the closest approximation of the position of the fish and always populated:

```{r}
occ_detections %<>% mutate(decimalLatitude = sprintf("%.7f", round(deployment_lat, digits = 7)))
```

Check `decimalLatitude` is not `NA` (should be 0):

```{r}
occ_detections %>%
  filter(is.na(decimalLatitude)) %>%
  nrow()
```

### decimalLongitude

```{r}
occ_detections %<>% mutate(decimalLongitude = sprintf("%.7f", round(deployment_long, digits = 7)))
```

### geodeticDatum

```{r}
occ_detections %<>% mutate(geodeticDatum = "WGS84")
```

### coordinateUncertaintyInMeters

```{r}
# Depends on area: sea / Westerscheldt: 200m on average, 500m extreme, while Albertkanaal: 2km
# TODO: on which field should this be based?
# network_project_project?
# animal_project_project?
# detections_receiver?
```

### georeferenceSources

```{r}
occ_detections %<>% mutate(georeferenceSources = "GPS") # TODO: not always GPS, maybe drop term
```

### scientificName

```{r}
occ_detections %<>% mutate(scientificName = animals_scientific_name)
```

Show unique values:

```{r}
occ_detections %>%
  select(animals_scientific_name) %>%
  group_by_all() %>%
  summarize(records = n())
```

Filter out records with `Sync tag` as scientific name:

```{r}
occ_detections %<>% filter(scientificName != "Sync tag")
```

Number of remaining occurrences:

```{r}
nrow(occ_detections)
```

### kingdom

```{r}
occ_detections %<>% mutate(kingdom = "Animalia")
```

Some other higher classication terms could be populated, but with the limited number of species it's not really useful as extra filters.

### taxonRank

```{r}
occ_detections %<>% mutate(taxonRank = "species") # TODO: all species?
```

## Terms considered but not included in mapping

### locality

There is location information in `deployments_location_name`, but these raw Dutch location names are not very useful, so we don't include this information:

```{r}
raw_detections %>%
  select(deployments_location_name) %>%
  unique() %>%
  head()
```

### minimumDepthInMeters

Pressure tags collect depth information, but that won't be available in the source data until VRL imports are supported.

### vernacularName

```{r}
raw_detections %>%
  select(animal_common_name) %>%
  unique() %>%
  head()
```

## Post-processing

Remove the original columns:

```{r}
occ_detections %<>% select(-one_of(raw_detections_colnames))
```

Preview data:

```{r}
occ_detections %>% head()
```

Save to CSV:

```{r}
write.csv(occ_detections, file = dwc_occ_detections_file, na = "", row.names = FALSE, fileEncoding = "UTF-8")
```
