---
title: "Darwin Core mapping"
subtitle: "ETN - LifeWatch.be acoustic telemetry data of fish"
author:
- Lien Reysehove
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
#  pdf_document:
#    df_print: kable
#    number_sections: yes
#    toc: yes
#    toc_depth: 3
---


# Setup 

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Install packages

```{r}
library(tidyverse)      # To do data science
library(magrittr)       # To use %<>% pipes
library(here)           # To find files
library(janitor)        # To clean input data
library(readxl)         # To read Excel files
library(digest)         # To generate hashes
library(rgbif)          # To use GBIF services
library(etn)
```

# Read source data

Connect to database:

```{r}
my_con <- connect_to_etn("damiano.oldoni@inbo.be", "1zFhM2Ic")
```

Load animal data:

```{r}
input_data <- get_animals(my_con)
tags    <- get_transmitters(my_con)
```

Save as input file

```{r}
write_csv(input_data, here("data", "processed", "raw.csv"), na = "")
```

Filter on requested animal projects:

```{r}
input_data %>% 
  group_by(projectname) %>% 
  summarise(records = n())
```

```{r}
input_data %<>% 
  filter(projectname %in% c("2011 Rivierprik",
                            "2012 Leopoldkanaal",
                            "2013 Albertkanaal",
                            "2014 Demer",
                            "2015 Dijle",
                            "2015 PhD Verhelst",
                            "Homarus",
                            "PhD Jan Reubens"))

# Exclude data form `scientific_name`= `Sync tag` in `projectname` = `2013 Albertkanaal` and
# for `scientific_name` = `Gadus morhua` in `projectname` = `2015 PhD Verhelst`
input_data <-
  input_data %>% 
    filter(!(projectname == "2013 Albertkanaal" & scientific_name == "Sync tag")) %>%
    filter(!(projectname == "2015 PhD Verhelst" & scientific_name == "Gadus morhua")) %>% 
    filter(!(projectname == "PhD Jan Reubens" & scientific_name == "Sentinel"))
```

# Process source data

## Tidy data

Clean data somewhat:

```{r}
input_data %<>% remove_empty("rows")
```

## Scientific names

Use the [GBIF nameparser](https://www.gbif.org/tools/name-parser) to retrieve nomenclatural information for the scientific names in the checklist:

```{r}
parsed_names <- input_data %>%
  distinct(scientific_name) %>%
  pull() %>% # Create vector from dataframe
  parsenames() # An rgbif function
```

Show scientific names with nomenclatural issues, i.e. not of `type = SCIENTIFIC` or that could not be fully parsed. Note: these are not necessarily incorrect.

```{r}
parsed_names %>%
  select(scientificname, type, parsed, parsedpartially, rankmarker) %>%
  filter(!(type == "SCIENTIFIC" & parsed == "TRUE" & parsedpartially == "FALSE"))
```

# Capture Event

# Surgery Event

# Release Event

